\chapter*{\uppercase{ABSTRACT}}
\addcontentsline{toc}{section}{\bfseries Abstract}

\begin{spacing}{1.5}

Competitive examination preparation requires aspirants to practice a large number of questions with a strong focus on topic-wise coverage and performance analysis. However, organizing large question banks and generating personalized test series manually is time-consuming, error-prone, and lacks systematic feedback. Most existing platforms provide generic mock tests with limited flexibility in test creation and minimal insight into topic-wise performance.

This project presents an Intelligent Competitive Exam Platform with Personalized Test Series that enables structured question management and focused practice. The system accepts Excel-based question input, converts it into a standardized JSON format, and automatically assigns topic and subtopic labels using NLP-based sentence embedding techniques. A hybrid classification approach ensures accurate tagging by combining semantic similarity methods with intelligent fallback resolution for ambiguous cases.

The platform allows users to generate customized test series by selecting one or more topics and specifying the number of questions per topic, while subtopic-level distribution is handled automatically. Time configuration is optional, with a default allocation of one minute per question. In addition, the system provides performance analysis with SWOT-based feedback, offering clear insights into strengths, weaknesses, opportunities, and areas for improvement based on topic-wise accuracy and test history.

The proposed system enhances preparation efficiency by enabling structured practice, flexible test generation, and meaningful performance feedback, thereby addressing key limitations of traditional mock-test platforms.

\end{sloppypar}
\end{spacing}

